---
title: "Assignment_2_Markdown"
author: "Kathleen Scopis"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# STEP 1

install packages and load in data

```{r warning=FALSE, message=FALSE, cache=FALSE}
options(scipen=999)

#install packages

#install.packages(c("sp", "rgdal", "rgeos", "spdep", "spgwr", "tmap", "spatialreg", "lmtest", "whitestrap", "tseries"))
                 
library(sp)
library(rgdal)
library(rgeos)
library(spdep)
library(spgwr)
library(tmap)
library(spatialreg)
library(whitestrap)
library(lmtest)
library(tseries)

pallete1 <- c("#2596be","#76b5c5","#abdbe3") #blues
pallete2 <- c("#e28743", "#eab676", "#eeeee4") #oranges
pallete3 <- c("#873e23", "#e28743", "#eeeee4") #browns


#read data
shp <- readOGR('C:\\Users\\kathl\\OneDrive\\Desktop\\Spatial Statistics\\Assignment2\\Regression Data')
names(shp@data)


```


#1.a
present histograms for NBELPOV100, then its log transformation (after adding +1)

```{r warning=FALSE, message=FALSE, cache=FALSE}

#raw variable
par(oma=c(0,0,2,0)) 
par(mfrow=c(1,3)) 
hist(shp@data$NBelPov100, breaks = 40, col = "#e28743",
     main = "Number of households with 
     income below 100% poverty",
     xlab = "NBelPov100")

#log transformation +1
shp@data$Log_NBelPov100 <- log(shp@data$NBelPov100 + 1)
shp@data$Log_MedHval <- log(shp@data$MEDHVAL + 1)

#histograms
par(oma=c(0,0,2,0)) 
par(mfrow=c(1,3)) 
hist(shp@data$Log_NBelPov100, breaks = 50,
     main = "Log Transformation of the
     # of Households Below Poverty",
     xlab = "Log_NBelPov100",
     col = "#eab676")

```


#1.b

introducing queen neighbor relationships

```{r warning=FALSE, message=FALSE, cache=FALSE}

queen<-poly2nb(shp, row.names=shp$POLY_ID)
summary(queen)

#plot links
plot(shp, col='grey90', lwd=2)
xy<-coordinates(shp)
par(mfrow=c(1,1)) 
plot(queen, xy, col='#d74821', lwd=.1, add=TRUE)
title(main='Contiguous Queen Neighbors')
```


it is observed that two regions only have one neighbor, while one region has as many as 27 direct queen neighbors - wow! 
I want to go trick-or-treating with them!

```{r warning=FALSE, message=FALSE, cache=FALSE}

#see which region has only one neighbor
smallestnbcard<-card(queen) #extract neighbor matrix
smallestnb<-which(smallestnbcard == min(smallestnbcard)) #extract block groups with smallest number of neighbors
fg<-rep('grey90', length(smallestnbcard))
fg[smallestnb]<-'#873e23' #color block groups dark orange
fg[queen[[smallestnb[1]]]]<-'#e28743' #color neighboring blocks light orange
fg[queen[[smallestnb[2]]]]<-'#e28743'
fg[queen[[smallestnb[3]]]]<-'#e28743'
fg[queen[[smallestnb[4]]]]<-'#e28743'
plot(shp, col=fg)
title(main='Regions with only 1 neighbor')

#see which region has most neighbors
largestnbcard<-card(queen)
largestnb<-which(largestnbcard == max(largestnbcard))
fg1<-rep('grey90', length(largestnbcard))
fg1[largestnb]<-'#873e23'
fg1[queen[[largestnb]]]<-'#e28743'
plot(shp, col=fg1)
title(main='Region with 27 neighbors')
```

#1.c

calculate global moran's i using queen weight matrix
it is observed that the global moran's i for Log_MedHval is ~0.79.

```{r warning=FALSE, message=FALSE, cache=FALSE}
queenlist<-nb2listw(queen, style = 'W')
moran(shp$Log_MedHval, queenlist, n=length(queenlist$neighbours), S0=Szero(queenlist))$`I` 

moranMC<-moran.mc(shp$LNMEDHHINC, queenlist, nsim=999, alternative="two.sided")  #We use 999 permutations
moranMC

moranMCres<-moranMC$res
hist(moranMCres, freq=10000000, nclass=100,
     col = "#e28743" )   #Draws distribution of Moran's I's calculated from randomly permuted values
# Here, we draw a red vertical line at the observed value of our Moran's I
abline(v=moran(shp$LNMEDHHINC, queenlist, n=length(queenlist$neighbours), S0=Szero(queenlist))$`I`, col='#c63112')

moran.plot(shp$LNMEDHHINC, queenlist, xlab = "LNMEDHHINC") 
```


#1.d
run local moran's I (LISA - Local Indicator of Spatial Association) analysis using Queen weight matrix.
is there spatial autocorrelation?  let's check the map...
we have also included a spatial association map, which projects areas that have high/high, high/low, low/low, and low/high relationships to their neighboring regions.

```{r warning=FALSE, message=FALSE, cache=FALSE}
#Run local moran's I (LISA) 
LISA<-localmoran(shp$Log_MedHval, queenlist)
head(LISA)
df.LISA <-cbind(shp, as.data.frame(LISA))



moranSig.plot<-function(df,listw, title){
  local<-localmoran(x=df$Log_MedHval, listw=listw, zero.policy = FALSE)
  moran.map<-cbind(df, local)
  #Here, col='Pr.z....E.Ii..' is the name of the column in the dataframe df.LISA that we're trying to plot. This variable name might change based on the version of the package.
  tm<-tm_shape(moran.map)+
    tm_borders(col='white')+
    tm_fill(style='fixed', col='Pr.z....E.Ii..', breaks=c(0,0.001, 0.01, 0.05, 1), title= 'p-value', palette = '-BuPu')+
    tm_layout(frame = FALSE, title = title)
  print(tm)
}
moranSig.plot(df.LISA, queenlist, 'p-value')


#spatial association map 
hl.plot<-function(df, listw){
  local<-localmoran(x=df$Log_MedHval, listw=listw, zero.policy = FALSE)
  quadrant<-vector(mode='numeric', length=323)
  m.prop<-df$Log_MedHval - mean(df$Log_MedHval)
  m.local<-local[,1]-mean(local[,1])
  signif<-0.05
  quadrant[m.prop >0 & m.local>0]<-4 #high MEDHHINC, high clustering
  quadrant[m.prop <0 & m.local<0]<-1 #low MEDHHINC, low clustering
  quadrant[m.prop <0 & m.local>0]<-2 #low MEDHINC, high clustering
  quadrant[m.prop >0 & m.local<0]<-3 #high MEDHHINC, low clustering
  quadrant[local[,5]>signif]<-0
  
  brks <- c(0,1,2,3,4)
  colors <- c("grey","light blue",'blue','pink',"red")
  plot<-plot(shp,border="gray90",lwd=1.0,col=colors[findInterval(quadrant,brks,all.inside=FALSE)])
}

hl.plot(shp, queenlist)
legend("bottomright",legend=c("insignificant","low-high","low-low","high-low","high-high"),
       fill=c("grey", "light blue", "blue", "pink", "red"),bty="n", cex = .8)

```

#1.e
OLS Regression

```{r warning=FALSE, message=FALSE, cache=FALSE}
reg<-lm(formula=Log_MedHval ~ Log_NBelPov100 + PCTBACHMOR + PCTSINGLES + PCTVACANT, data=shp@data)
summary(reg)



##Prints the log likelihood
#logLik(reg)                  
##Prints the results of the Breusch-Pagan Test to assess whether heteroscedasticity is present (package: lmtest)
#bptest(reg, studentize=FALSE)
##Prints the results of the Koenker-Bassett Test (also known as the Studentized Breusch-Pagan Test) to assess #whether heteroscedasticity is present (package: lmtest)
#bptest(reg)       
##Prints the results of the White Test to assess whether heteroscedasticity is present (package: whitestrap)
#white_test(reg)   
##Prints the results of the Jarque-Bera Test to assess whether residuals are normal (package: tseries)
#jarque.bera.test(reg$residuals)
```

#1.f CHECK

not sure if this section is correct - are these the proper weighted residuals?
line 223

Standardized OLS Residuals (OLS Model residuals divided by an estimate of their standard deviation) 
  There appears to be some autocorrelation.

```{r warning=FALSE, message=FALSE, cache=FALSE}
standardised<-rstandard(reg)
resnb<-sapply(queen, function(x) mean(standardised[x]))

shp@data$standardised <- standardised    #creating a new variable in the shapefile shp.
OLS.Residuals.Map<-tm_shape(shp)+
  tm_fill(col='standardised', style='quantile', title='Standardized OLS Residuals', 
          palette ='Blues')+
  tm_layout(frame=FALSE, title = 'Standardised OLS Residuals')
OLS.Residuals.Map

```

Regress the standardized OLS residuals on the spatial lag of the OLS residuals.

```{r warning=FALSE, message=FALSE, cache=FALSE}
#Regressing residuals on their nearest neighbors.
res.lm <- lm(formula=standardised ~ resnb)
summary(res.lm)
```


#1.g CHECK
Calculate Moran's I statistic and a pseudo p-value.

```{r warning=FALSE, message=FALSE, cache=FALSE}
moran.mc(standardised, queenlist, 999, alternative="two.sided")
moran.plot(standardised, queenlist)
```

#1.h  CHECK - IS THIS CORRECT? (MIXED UP SPATIAL LAG AND SPATIAL ERROR) + ADD MAP
Moran's I of Spatial Lag Model residuals. (Again, looking at the map of the Spatial Error Model residuals is an exercise left for the student.)

```{r warning=FALSE, message=FALSE, cache=FALSE}
reslag<-lagreg$residuals
lagMoranMc<-moran.mc(reslag, queenlist,999, alternative="two.sided")
lagMoranMc

# Add map here

```


#1.i  CHECK - are the spatial lag residual regressions correct?  is this the right way to run 999 permutations?
Spatial Lag Regression

```{r warning=FALSE, message=FALSE, cache=FALSE}


lagreg<-lagsarlm(formula=Log_MedHval ~ Log_NBelPov100 + PCTBACHMOR + PCTSINGLES + PCTVACANT, data=shp@data,
               queenlist)
summary(lagreg)
LR.Sarlm(lagreg, reg) #Here lagreg is the SL output; reg is the OLS output
#Prints the results of the Breusch-Pagan Test to assess whether heteroscedasticity is present (package: lmtest)
bptest.Sarlm(lagreg, studentize=FALSE)
#Prints the results of the Koenker-Bassett Test (also known as the Studentized Breusch-Pagan Test) to assess whether heteroscedasticity is present (package: lmtest)
bptest.Sarlm(lagreg)       
#Prints the results of the Jarque-Bera Test to assess whether residuals are normal (package: tseries)
jarque.bera.test(lagreg$residuals)



# 999 permutations - is this missing something?  is this correct?

reslag<-lagreg$residuals
lagMoranMc<-moran.mc(reslag, queenlist,999, alternative="two.sided")
lagMoranMc

```


1.j CHECK - MIXED UP SPATIAL LAG AND SPATIAL ERROR
Spatial Error Regression


```{r warning=FALSE, message=FALSE, cache=FALSE}
errreg<-errorsarlm(formula=Log_MedHval ~ Log_NBelPov100 + PCTBACHMOR + PCTSINGLES + PCTVACANT, data=shp@data,
                   queenlist)
reserr<-residuals(errreg)
errresnb<-sapply(queen, function(x) mean(reserr[x]))
summary(errreg)
LR.Sarlm(errreg, reg)
#Prints the results of the Breusch-Pagan Test to assess whether heteroscedasticity is present (package: lmtest)
bptest.Sarlm(errreg, studentize=FALSE)
#Prints the results of the Koenker-Bassett Test (also known as the Studentized Breusch-Pagan Test) to assess whether heteroscedasticity is present (package: lmtest)
bptest.Sarlm(errreg)       
#Prints the results of the Jarque-Bera Test to assess whether residuals are normal (package: tseries)
jarque.bera.test(errreg$residuals)



errMoranMc<-moran.mc(reserr, queenlist, 999, alternative="two.sided")
errMoranMc
```


#2
GWR - Geographically Weighted Regression


```{r warning=FALSE, message=FALSE, cache=FALSE}
#Setting an adaptive bandwidth
bw<-gwr.sel(formula=Log_MedHval ~ Log_NBelPov100 + PCTBACHMOR + PCTSINGLES + PCTVACANT, 
            data=shp@data,
            method = "aic",
            adapt = TRUE)
bw



#setting a fixed bandwidth
bw_fixed<-gwr.sel(formula=Log_MedHval ~ Log_NBelPov100 + PCTBACHMOR + PCTSINGLES + PCTVACANT, 
            data=shp@data,
            method = "aic",
            adapt = FALSE)
bw_fixed


#Adaptive Bandwidth

gwrmodel<-gwr(formula=LNMEDHHINC~LNMEDHVAL+PCTVACANT,
              data=shp,
              adapt = bw, #adaptive bandwidth determined by proportion of observations accounted for
              gweight=gwr.Gauss,
              se.fit=TRUE, #to return local standard errors
              hatmatrix = TRUE)
gwrmodel


#Fixed Bandwidth

gwrmodel_fixed<-gwr(formula=LNMEDHHINC~LNMEDHVAL+PCTVACANT,
              data=shp,
              bandwidth = bw_fixed, #fixed bandwidth
              gweight=gwr.Gauss,
              se.fit=TRUE, #to return local standard errors
              hatmatrix = TRUE)
gwrmodel_fixed


#GWR Output
summary(gwrmodel$SDF)

############################  QUESTION: Do we plug in the predictive variables here?

#map standardized coefficients
gwrresults<-as.data.frame(gwrmodel$SDF)
shp$coefLNMEDHVALst<-gwrresults$LNMEDHVAL/gwrresults$LNMEDHVAL_se
shp$coefPCTVACANTst<-gwrresults$PCTVACANT/gwrresults$PCTVACANT_se

shp$gwrE<-gwrresults$gwr.e
shp$localR2<-gwrresults$localR2

coefLNMEDHVAL<-tm_shape(shp)+
  tm_fill(col='coefLNMEDHVALst', breaks=c(-Inf, -6, -4, -2, 0, 2, 4, 6, Inf), title='Standardized coefficient of LNMEDHVAL', 
          palette ='-RdBu')+
  tm_layout(frame=FALSE, title = 'Median House Value (Log)')

coefPCTVACANT<-tm_shape(shp)+
  tm_fill(col='coefPCTVACANTst', breaks=c(-Inf, -6, -4, -2, 0, 2, 4, 6, Inf), title='Standardized coefficient of PCTVACANT', 
          palette='-RdBu')+
  tm_layout(frame=FALSE, title = 'Percentage of Housing Vacant')

tmap_arrange(coefLNMEDHVAL, coefPCTVACANT, ncol=2)

tm_shape(shp)+
  tm_fill(col='localR2',  breaks=c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7), n=5, palette = 'Blues')+
  tm_layout(frame=FALSE)


```












